---
title: 'Exercise #1'
author: "Serhan Bayram"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Temperature Conversion

The function:

```{r}
convert_temperature <- function(temperature,F_to_C){
  if (F_to_C == FALSE) {
    new_temperature = (temperature * 9 / 5) + 32
  }
  else{
    new_temperature = (temperature - 32) * 5 / 9
  }
  return(new_temperature)
}
```

Execution to convert from F to C:

```{r}
convert_temperature(temperature = 86,F_to_C = TRUE)
```

Execution to convert from C to F:

```{r}
convert_temperature(temperature = 30,F_to_C = FALSE)
```

### Future Value of an Investment

The function:

```{r}
calculate_future_value <- function(investment, interest, duration_in_years){
  future_val = investment * (1 + interest)^duration_in_years
  return(future_val)
}
```

Execution for 100 units of investments 7% interest rate over 5 years:

```{r}
calculate_future_value(investment = 100, interest = 0.07, duration_in_years = 5)
```

### Generating Color Hex Codes

The function:

```{r}
generate_hex_code <- function(n){
  num_vals <- c(0:9)
  letter_vals <- c("a","b","c","d","e","f")
  all_vals <- append(num_vals, letter_vals)
  hex_code <- ""
  hex_array <- c()
  for (y in 1:n) {
    for (x in 1:6) {
      hex_code <- paste0(hex_code, sample(all_vals,1))
    }
    hex_code <- paste0("#",hex_code)
    hex_array[y] <- hex_code
    hex_code <- ""
  }
  return(hex_array)
}
```

Color hex code generation:

```{r}
generate_hex_code(n=3)
```

### Calculate Probability of Dice

The function:

```{r}
get_prob_dice <- function(six_count, throw_count){
  probability_rate <- dbinom(six_count, size=throw_count, prob=1/6) 
  return(probability_rate)
}
```

Probability of getting 3 sixes in 5 throws:

```{r}
get_prob_dice(3,5)
```

### Rock, Scissors, Paper

The function:

```{r}
rsp_game <- function(choice){
  possible_choices <- c("rock","scissors","paper")
  func_choice <- sample(1:3, 1)
  my_choice <- possible_choices[func_choice]
  if (my_choice == choice) {
    return("I choose the same. Tie!")
  }
  else if(my_choice == "rock" & choice == "scissors"){
    return("You lost, I choosed rock.")
  } 
  else if(my_choice == "rock" & choice == "paper"){
    return("You won, I choosed rock.")
  } 
  else if(my_choice == "scissors" & choice == "rock"){
    return("You won, I choosed scissors")
  } 
  else if(my_choice == "scissors" & choice == "paper"){
    return("You lost, I choosed scissors")
  } 
  else if(my_choice == "paper" & choice == "rock"){
    return("You lost, I choosed paper")
  } 
  else if(my_choice == "paper" & choice == "scissors"){
    return("You won, I choosed paper")
  } 
}
```

Sample game results:

```{r}
rsp_game("rock")
```

```{r}
rsp_game("scissors")
```

```{r}
rsp_game("paper")
```

Back to my [Progress Journal](https://pjournal.github.io/mef04-SBMEFBDA/)